Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models

### 2020-08-16 performance/performance.tex
Almost finished

@@

The variant total required samples for different runs under the same setting was most likely the result of different threshold evolve path. Median epsilon schedule and randomness in the sampling process resulted in different sequences of threshold for different runs, which means different repeated runs may have different convergency path \cite{threshold}. For example, Some runs temporarily converges to a local optimum, while the target epsilon in the next generation cannot possibly be reached if the new particles are still concentrated

will take huge amount of samples to move away from the local optimum

Moreover, we found that the execution time was not ideally proportional to the sampling tires in that period of time, for both different repeated runs and different generations in a single run. To illustrate this, a single run from $p=24$ was selected and visualised.

A possible explain to these phenomenons can be that the median epsilon schedule results in different threshold strategies due to the randomness in the sampling process, thus results in different approximate convergency paths for different runs. Some runs that are stuck in a local optimum and require smaller target epsilon will take huge amount of samples to move away from the local optimum. This have been observed when the distribution of posterior is plotted before and after jumping away from local modes (FIGURE and more explain). Also, time spent in each sample using sanme number of cores is not constant in our case, which might be a result of the parallel schedule; ideally a serial run using only one core will see a nearly constant time spent in every particle.

>>>>

\section{Discussions}

@@

\begin{figure}[ht]
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/local_modes.png}}
    \end{center}

    \caption[TODO]{Local modes. MORE DESCRIPTION}
    \label{fig:local_modes}
\end{figure}

The variant total required samples for different runs under the same setting was most likely the result of different threshold evolve path. Median epsilon schedule and randomness in the sampling process resulted in different sequences of threshold for different runs, which means different repeated runs may have different convergency path \cite{threshold}. For example, Some runs temporarily converges to a local optimum, while the target epsilon in the next generation cannot possibly be reached if the new particles are still concentrated in the local optimum. In this case, huge amount of sample will be drawn to find enough accepted particles, and new generation will be able to jump out of the local optimum and have different concentrations.

To illustrate this, Two independent run, namely Run A and Run B, were chosen from the experiment of $p=24$ and visualised in Figure \ref{fig:local_modes} and Figure \ref{fig:local_para}. Thess two run used identical options of the algorithm. Figure \ref{fig:local_modes} (a) shows the comparison of the required samples, where Run A converged to a local mode, and in population 11 took much more sample tires to move away from the local optimum (Figure \ref{fig:local_modes} (b)). Figure \ref{fig:local_para} compares the posterior density distribution before (t=9, cyan) and after (t=13, red) the jumping away: significant movement of concentrations in posterior density distributions are observed, especially in parameters from ODE 1, 2 and 4. An interning movement is that there are two concentrations in $\kappa_{\Phi\beta}$ and $\mu_\Phi$ when t=9, and the majority of particles are concentrated in the second concentration with high parameter values; when t=13, the second concentration distribution disappears and all particles are concentrated near zero.

\begin{figure}[t]
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/local_para.png}}
    \end{center}

    \caption[TODO]{Parameter density distribution. MORE DESCRIPTION}
    \label{fig:local_para}
\end{figure}

Moreover, we found that the execution time was not ideally proportional to the sampling tires in that period of time, for both different repeated runs and different generations in a single run. From the same example, Figure \ref{fig:local_modes} (b) and (c) show the non-linear relationship between execution time and sample tries in different generations. Apart from 1st and 2nd generation, generation 3 to 20 show a general linear relationship i.e. $T_t\propto n_t$, where $t$ is the population index, $T_t$ is the execution time of generation $t$ and $n_t$ is the required samples in generation $t$. However, this relationship does not hold for the first two generation. Generation 1 is to draw sample from prior, but it is done with high efficiency as only small amount of time are taken; generation 2 takes unusual large amount of the, but it does not require much more samples. The abnormal execution time of generation 1 and two are observed in all other runs, where for $p=24$ generation 1 usually takes c. 16s and generation 2 takes c. 1600s. This is believed to be related to the initialisation of the algorithm.

Additionally, by observing CPU frequency and load curve in-time while running the program, we found that the parallelised part is limited the the sampling, calculation and criterion test process; some reduction calculations and preparation work for the next generation (e.g. calculation of the new epsilon values, normalisation of data, fit of multivariate kernel) are serial. Database I/O is also serial.

% A possible explain to these phenomenons can be that the median epsilon schedule results in different threshold strategies due to the randomness in the sampling process, thus results in different approximate convergency paths for different runs. Some runs that are stuck in a local optimum and require smaller target epsilon will take huge amount of samples to move away from the local optimum. This have been observed when the distribution of posterior is plotted before and after jumping away from local modes (FIGURE and more explain). Also, time spent in each sample using sanme number of cores is not constant in our case, which might be a result of the parallel schedule; ideally a serial run using only one core will see a nearly constant time spent in every particle.
